---
title: "Baseball as a Motivating Example"
author: "Sajini Arumugam"
output: html_document
weight: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Assessment: Baseball as a Motivating Example. Linear regression

#### **Question 1**  
What is the application of statistics and data science to baseball called?
- Moneyball
- Sabermetrics    [*]
- The “Oakland A’s Approach”
- There is no specific name for this; it’s just data science.

#### **Question 2**  

Which of the following outcomes is not included in the batting average?

- A home run
- A base on balls   [*]
- An out
- A single

#### **Question 3**  

Why do we consider team statistics as well as individual player statistics?
- The success of any individual player also depends on the strength of their team.   [*]
- Team statistics can be easier to calculate.
- The ultimate goal of sabermetrics is to rank teams, not players.


#### **Question 4**  

You want to know whether teams with more at-bats per game have more runs per game.

What R code below correctly makes a scatter plot for this relationship?

```
Teams %>% filter(yearID %in% 1961:2001 ) %>%
    ggplot(aes(AB, R)) + 
    geom_point(alpha = 0.5)
```

```
[*]
Teams %>% filter(yearID %in% 1961:2001 ) %>%
    mutate(AB_per_game = AB/G, R_per_game = R/G) %>%
    ggplot(aes(AB_per_game, R_per_game)) + 
    geom_point(alpha = 0.5)
```

```
Teams %>% filter(yearID %in% 1961:2001 ) %>%
    mutate(AB_per_game = AB/G, R_per_game = R/G) %>%
    ggplot(aes(AB_per_game, R_per_game)) + 
    geom_line()
```

```
Teams %>% filter(yearID %in% 1961:2001 ) %>%
    mutate(AB_per_game = AB/G, R_per_game = R/G) %>%
    ggplot(aes(R_per_game, AB_per_game)) + 
    geom_point()
```

#### **Question 5**  

What does the variable “SOA” stand for in the Teams table?


- sacrifice out
- slides or attempts
- strikeouts by pitchers    [*]
- accumulated singles

#### **Question 6**  

Load the Lahman library. Filter the Teams data frame to include years from 1961 to 2001. Make a scatterplot of runs per game versus at bats (AB) per game.

Which of the following is true?

- There is no clear relationship between runs and at bats per game.
- As the number of at bats per game increases, the number of runs per game tends to increase.    [*]
- As the number of at bats per game increases, the number of runs per game tends to decrease.


#### **Question 7**  

Use the filtered Teams data frame from Question 6. Make a scatterplot of win rate (number of wins per game) versus number of fielding errors (E) per game.

Which of the following is true?

- There is no clear relationship between win rate and errors per game.
- As the number of errors per game increases, the win rate tends to increase.
- As the number of errors per game increases, the win rate tends to decrease.   [*]

#### **Question 8**  

Use the filtered Teams data frame from Question 6. Make a scatterplot of triples (X3B) per game versus doubles (X2B) per game.

Which of the following is true?

- There is no clear relationship between doubles per game and triples per game.    [*]
- As the number of doubles per game increases, the number of triples per game tends to increase.
- As the number of doubles per game increases, the number of triples per game tends to decrease.


### Assessment: Correlation

#### **Question 1**  

While studying heredity, Francis Galton developed what important statistical concept?

- Standard deviation
- Normal distribution
- Correlation   [*]
- Probability


#### **Question 2**  

The correlation coefficient is a summary of what?

- The trend between two variables   [*]
- The dispersion of a variable
- The central tendency of a variable
- The distribution of a variable

#### **Question 3**  

Below is a scatter plot showing the relationship between two variables, x and y.

Scatter plot of relationship between x (plotted on the x-axis) and y (plotted on the y-axis). y-axis values range from -3 to 3; x-axis values range from -3 to 3. Points are fairly well distributed in a tight band with a range from approximately (-2, 2) to (3, -3).
From this figure, the correlation between x and y appears to be about:

- -0.9    [*]
- -0.2
- 0.9
- 2


#### **Question 4**  

Instead of running a Monte Carlo simulation with a sample size of 25 from the 179 father-son pairs described in the videos, we now run our simulation with a sample size of 50.

Would you expect the expected value of our sample correlation to increase, decrease, or stay approximately the same?

- Increase
- Decrease
- Stay approximately the same   [*]

#### **Question 5**  

Instead of running a Monte Carlo simulation with a sample size of 25 from the 179 father-son pairs described in the videos, we now run our simulation with a sample size of 50.

Would you expect the standard error of our sample correlation to increase, decrease, or stay approximately the same?

- Increase
- Decrease    [*]
- Stay approximately the same


#### **Question 6**  

If X and Y are completely independent, what do you expect the value of the correlation coefficient to be?

- -1
- -0.5
- 0   [*]
- 0.5
- 1
- Not enough information to answer the question


#### **Question 7**  

Load the Lahman library. Filter the Teams data frame to include years from 1961 to 2001.

What is the correlation coefficient between number of runs per game and number of at bats per game?

```{r}
library(Lahman)
library(dplyr)
library(ggplot2)
Teams_small <- Teams %>% filter(yearID %in% 1961:2001)
cor(Teams_small$R/Teams_small$G, Teams_small$AB/Teams_small$G)
```


#### **Question 8**  

Use the filtered Teams data frame from Question 7.

What is the correlation coefficient between win rate (number of wins per game) and number of errors per game?

```{r}
cor(Teams_small$W/Teams_small$G, Teams_small$E/Teams_small$G)
```


#### **Question 9**  

Use the filtered Teams data frame from Question 7.

What is the correlation coefficient between doubles (X2B) per game and triples (X3B) per game?

```{r}
cor(Teams_small$X2B/Teams_small$G, Teams_small$X3B/Teams_small$G)
```


### Assessment: Stratification and Variance Explained, Part 1

#### **Question 1**  

- Slope = (correlation coefficient of son and father heights) * (standard deviation of sons’ heights / standard deviation of fathers’ heights)[*]
- Slope = (correlation coefficient of son and father heights) * (standard deviation of fathers’ heights / standard deviation of sons’ heights)
- Slope = (correlation coefficient of son and father heights) / (standard deviation of sons’ heights * standard deviation of fathers’ heights)
- Slope = (mean height of fathers) - (correlation coefficient of son and father heights * mean height of sons).


Question 2

Why does the regression line simplify to a line with intercept zero and slope  ρ  when we standardize our x and y variables?
Try the simplification on your own first!

- When we standardize variables, both x and y will have a mean of one and a standard deviation of zero. When you substitute this into the formula for the regression line, the terms cancel out until we have the following equation:  yi=ρxi .
- When we standardize variables, both x and y will have a mean of zero and a standard deviation of one. When you substitute this into the formula for the regression line, the terms cancel out until we have the following equation:  yi=ρxi .   [*]
- When we standardize variables, both x and y will have a mean of zero and a standard deviation of one. When you substitute this into the formula for the regression line, the terms cancel out until we have the following equation:  yi=ρ+xi .


Question 3

What is a limitation of calculating conditional means?
Select ALL that apply.

- Each stratum we condition on (e.g., a specific father’s height) may not have many data points.[*]
- Because there are limited data points for each stratum, our average values have large standard errors.[*]
- Conditional means are less stable than a regression line.[*]
- Conditional means are a useful theoretical tool but cannot be calculated.


Question 4

A regression line is the best prediction of Y given we know the value of X when:

- X and Y follow a bivariate normal distribution.[*]
- Both X and Y are normally distributed.
- Both X and Y have been standardized.
- There are at least 25 X-Y pairs.


Question 5

Which one of the following scatterplots depicts an x and y distribution that is NOT well-approximated by the bivariate normal distribution?

[*] The v-shaped distribution of points from the first plot means that the x and y variables do not follow a bivariate normal distribution.

When a pair of random variables is approximated by a bivariate normal, the scatter plot looks like an oval (as in the 2nd, 3rd, and 4th plots) - it is okay if the oval is very round (as in the 3rd plot) or long and thin (as in the 4th plot).


Question 6

We previously calculated that the correlation coefficient  ρ  between fathers’ and sons’ heights is 0.5.

Given this, what percent of the variation in sons’ heights is explained by fathers’ heights?

- 0%
- 25%   [*]
- 50%
- 75%


Question 7

Suppose the correlation between father and son’s height is 0.5, the standard deviation of fathers’ heights is 2 inches, and the standard deviation of sons’ heights is 3 inches.

Given a one inch increase in a father’s height, what is the predicted change in the son’s height?

- 0.333
- 0.5
- 0.667
- 0.75    [*]


```{r}
set.seed(1989, sample.kind="Rounding")
library(HistData)
data("GaltonFamilies")

female_heights <- GaltonFamilies%>%    
    filter(gender == "female") %>%    
    group_by(family) %>%    
    sample_n(1) %>%    
    ungroup() %>%    
    select(mother, childHeight) %>%    
    rename(daughter = childHeight)
```

