---
title: "Capstone - Movielens"
author: "Sajini Arumugam"
output: html_document
weight : 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(data.table)) install.packages("data.table", repos = "http://cran.us.r-project.org")

dl <- tempfile()
download.file("http://files.grouplens.org/datasets/movielens/ml-10m.zip", dl)

ratings <- fread(text = gsub("::", "\t", readLines(unzip(dl, "ml-10M100K/ratings.dat"))),
                 col.names = c("userId", "movieId", "rating", "timestamp"))

movies <- str_split_fixed(readLines(unzip(dl, "ml-10M100K/movies.dat")), "\\::", 3)
colnames(movies) <- c("movieId", "title", "genres")

movies <- as.data.frame(movies) %>% mutate(movieId = as.numeric(levels(movieId))[movieId],
                                           title = as.character(title),
                                           genres = as.character(genres))


movielens <- left_join(ratings, movies, by = "movieId")

set.seed(1, sample.kind="Rounding")
# if using R 3.5 or earlier, use `set.seed(1)` instead
test_index <- createDataPartition(y = movielens$rating, times = 1, p = 0.1, list = FALSE)
edx <- movielens[-test_index,]
temp <- movielens[test_index,]

# Make sure userId and movieId in validation set are also in edx set
validation <- temp %>% 
  semi_join(edx, by = "movieId") %>%
  semi_join(edx, by = "userId")

# Add rows removed from validation set back into edx set
removed <- anti_join(temp, validation)
edx <- rbind(edx, removed)

rm(dl, ratings, movies, test_index, temp, movielens, removed)


```


### Quiz: MovieLens Dataset

All of the questions in this quiz involve working with the datasets you've created using the code we provided.

#### **Question 1**  

How many rows and columns are there in the edx dataset?

Number of rows:

```{r}
dim(edx)
```

#### **Question 2**  

How many zeros were given as ratings in the edx dataset?

```{r}
head(edx)

j <- edx %>% filter(rating == 3)
nrow(j)
```


#### **Question 3**  
How many different movies are in the edx dataset?

```{r}
n_distinct(edx$movieId)
```


#### **Question 4**  
How many different users are in the edx dataset?

```{r}
n_distinct(edx$userId)
```


#### **Question 5**  
How many movie ratings are in each of the following genres in the edx dataset?

```{r}
n_drama <- edx %>% separate_rows(genres, sep = "\\|")%>%
  group_by(genres)
```

#### **Question 6**  
Which movie has the greatest number of ratings?

```{r}
n_rating <- edx %>% group_by(movieId,title) %>%
  summarize(count = n())%>% arrange(desc(count))
n_rating
```


#### **Question 7**  
What are the five most given ratings in order from most to least?

```{r}
edx %>% group_by(rating)%>%summarize(count = n())%>%
  arrange(desc(count))
```


#### **Question 8**  
True or False: In general, half star ratings are less common than whole star ratings (e.g., there are fewer ratings of 3.5 than there are ratings of 3 or 4, etc.).

```{r}
edx %>% group_by(rating)%>%summarize(count = n())%>%
  arrange(desc(count))
```



