---
title: "01 - Introduction to Discrete Probability"
author: "Sajini Arumugam"
date: "7/15/2020"
output: html_document
weight : 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Assessment: Introduction to Discrete Probability**  

#### **Probability of cyan**  

One ball will be drawn at random from a box containing: 3 cyan balls, 5 magenta balls, and 7 yellow balls.

What is the probability that the ball will be cyan?
1/5

#### **Probability of not cyan**  

One ball will be drawn at random from a box containing: 3 cyan balls, 5 magenta balls, and 7 yellow balls.

What is the probability that the ball will not be cyan?
4/5

#### **Sampling without replacement**  

Instead of taking just one draw, consider taking two draws. You take the second draw without returning the first draw to the box. We call this sampling without replacement.

What is the probability that the first draw is cyan and that the second draw is not cyan?

(3/15 * 12/14)

#### **Sampling with replacement**  

Now repeat the experiment, but this time, after taking the first draw and recording the color, return it back to the box and shake the box. We call this sampling with replacement.

What is the probability that the first draw is cyan and that the second draw is not cyan?

(3/15 * 12/15)

***

#### **Exercise 1. Probability of cyan - generalized**  
In the edX exercises for this section, we calculated some probabilities by hand. Now we'll calculate those probabilities using R.

One ball will be drawn at random from a box containing: 3 cyan balls, 5 magenta balls, and 7 yellow balls.

What is the probability that the ball will be cyan?

```{r}
cyan <- 3
magenta <- 5
yellow <- 7

# Assign a variable `p` as the probability of choosing a cyan ball from the box
p<-cyan/(cyan+magenta+yellow)

# Print the variable `p` to the console
p
```

#### **Exercise 2. Probability of not cyan - generalized**  
We defined the variable p as the probability of choosing a cyan ball from a box containing: 3 cyan balls, 5 magenta balls, and 7 yellow balls.

What is the probability that the ball you draw from the box will NOT be cyan?
```{r}
# `p` is defined as the probability of choosing a cyan ball from a box containing: 3 cyan balls, 5 magenta balls, and 7 yellow balls.
# Using variable `p`, calculate the probability of choosing any ball that is not cyan from the box
1-p
```

#### **Exercise 3. Sampling without replacement - generalized**  
Instead of taking just one draw, consider taking two draws. You take the second draw without returning the first draw to the box. We call this sampling without replacement.

What is the probability that the first draw is cyan and that the second draw is not cyan?
```{r}
cyan <- 3
magenta <- 5
yellow <- 7

# The variable `p_1` is the probability of choosing a cyan ball from the box on the first draw.
p_1 <- cyan / (cyan + magenta + yellow)
j<-cyan+magenta+yellow
# Assign a variable `p_2` as the probability of not choosing a cyan ball on the second draw without replacement.
p_2<- 1-(cyan-1)/(j-1)

# Calculate the probability that the first draw is cyan and the second draw is not cyan using `p_1` and `p_2`.
p_1*p_2
```

#### **Exercise 4. Sampling with replacement - generalized**  
Now repeat the experiment, but this time, after taking the first draw and recording the color, return it back to the box and shake the box. We call this sampling with replacement.

What is the probability that the first draw is cyan and that the second draw is not cyan?

```{r}
cyan <- 3
magenta <- 5
yellow <- 7

# The variable 'p_1' is the probability of choosing a cyan ball from the box on the first draw.
p_1 <- cyan / (cyan + magenta + yellow)
j<-cyan+magenta+yellow
# Assign a variable 'p_2' as the probability of not choosing a cyan ball on the second draw with replacement.
p_2 <- 1-(cyan/j)
# Calculate the probability that the first draw is cyan and the second draw is not cyan using `p_1` and `p_2`.
p_1*p_2
```

### **1.2 Combinations and Permutations**  

### **DataCamp Assessment: Combinations and Permutations**  

#### **Exercise 1. Independence**  
Imagine you draw two balls from a box containing colored balls. You either replace the first ball before you draw the second or you leave the first ball out of the box when you draw the second ball.

Under which situation are the two draws independent of one another?

Remember that two events A and B are independent if Pr(A and B)=Pr(A)P(B).


- You don't replace the first ball before drawing the next. [*]
- You do replace the first ball before drawing the next.


#### **Exercise 2. Sampling with replacement**  
Say you’ve drawn 5 balls from the a box that has 3 cyan balls, 5 magenta balls, and 7 yellow balls, with replacement, and all have been yellow.

What is the probability that the next one is yellow?

```{r}
cyan <- 3
magenta <- 5
yellow <- 7
j<-cyan+magenta+yellow
# Assign the variable 'p_yellow' as the probability that a yellow ball is drawn from the box.
p_yellow <- yellow/j

# Using the variable 'p_yellow', calculate the probability of drawing a yellow ball on the sixth draw. Print this value to the console.
p_yellow

```

#### **Exercise 3. Rolling a die**  
If you roll a 6-sided die once, what is the probability of not seeing a 6? If you roll a 6-sided die six times, what is the probability of not seeing a 6 on any of those rolls?
```{r}
# Assign the variable 'p_no6' as the probability of not seeing a 6 on a single roll.
p_no6 <- 5/6

# Calculate the probability of not seeing a 6 on six rolls using `p_no6`. Print your result to the console: do not assign it to a variable.
p_no6^6
```

#### **Exercise 4. Probability the Celtics win a game**  
Two teams, say the Celtics and the Cavs, are playing a seven game series. The Cavs are a better team and have a 60% chance of winning each game.

What is the probability that the Celtics win at least one game? Remember that the Celtics must win one of the first four games, or the series will be over!

```{r}
# Assign the variable `p_cavs_win4` as the probability that the Cavs will win the first four games of the series.
p_cavs_win4<-0.6^4

# Using the variable `p_cavs_win4`, calculate the probability that the Celtics win at least one game in the first four games of the series.
1-p_cavs_win4
```

#### **Exercise 5. Monte Carlo simulation for Celtics winning a game**  
Create a Monte Carlo simulation to confirm your answer to the previous problem by estimating how frequently the Celtics win at least 1 of 4 games. Use B <- 10000 simulations.

The provided sample code simulates a single series of four random games, simulated_games.

```{r}
# This line of example code simulates four independent random games where the Celtics either lose or win. Copy this example code to use within the `replicate` function.
simulated_games <- sample(c("lose","win"), 4, replace = TRUE, prob = c(0.6, 0.4))

# The variable 'B' specifies the number of times we want the simulation to run. Let's run the Monte Carlo simulation 10,000 times.
B <- 10000

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling.
set.seed(1)

# Create an object called `celtic_wins` that replicates two steps for B iterations: (1) generating a random four-game series `simulated_games` using the example code, then (2) determining whether the simulated series contains at least one win for the Celtics.
celtic_wins<- replicate(B, {
simulated_games <- sample(c("lose","win"), 4, replace = TRUE, prob = c(0.6, 0.4))
any(simulated_games=="win")
})


# Calculate the frequency out of B iterations that the Celtics won at least one game. Print your answer to the console.
mean(celtic_wins)
```

## **DataCamp Assessment: The Addition Rule and Monty Hall**  

#### **Exercise 1. The Cavs and the Warriors**  
Two teams, say the Cavs and the Warriors, are playing a seven game championship series. The first to win four games wins the series. The teams are equally good, so they each have a 50-50 chance of winning each game.

If the Cavs lose the first game, what is the probability that they win the series?

```{r}
# Assign a variable 'n' as the number of remaining games.
n<-6

# Assign a variable `outcomes` as a vector of possible game outcomes, where 0 indicates a loss and 1 indicates a win for the Cavs.

outcomes<-c(0,1)
# Assign a variable `l` to a list of all possible outcomes in all remaining games. Use the `rep` function on `list(outcomes)` to create list of length `n`.
l<-list(0:1)

# Create a data frame named 'possibilities' that contains all combinations of possible outcomes for the remaining games.

possibilities<-expand.grid(rep(l,n))
# Create a vector named 'results' that indicates whether each row in the data frame 'possibilities' contains enough wins for the Cavs to win the series.

results<-rowSums(possibilities)>=4

# Calculate the proportion of 'results' in which the Cavs win the series. Print the outcome to the console.

mean(results)
```

#### **Exercise 2. The Cavs and the Warriors - Monte Carlo**  
Confirm the results of the previous question with a Monte Carlo simulation to estimate the probability of the Cavs winning the series after losing the first game.

```{r}
# The variable `B` specifies the number of times we want the simulation to run. Let's run the Monte Carlo simulation 10,000 times.
B <- 10000

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling.
set.seed(1)

```

#### **Exercise 3. A and B play a series - part 1**  
Two teams, A and B, are playing a seven series game series. Team A is better than team B and has a p>0.5 chance of winning each game.

```{r}
# Let's assign the variable 'p' as the vector of probabilities that team A will win.
p <- seq(0.5, 0.95, 0.025)

# Given a value 'p', the probability of winning the series for the underdog team B can be computed with the following function based on a Monte Carlo simulation:
prob_win <- function(p){
  B <- 10000
  result <- replicate(B, {
    b_win <- sample(c(1,0), 7, replace = TRUE, prob = c(1-p, p))
    sum(b_win)>=4
    })
  mean(result)
}

# Apply the 'prob_win' function across the vector of probabilities that team A will win to determine the probability that team B will win. Call this object 'Pr'.

Pr<-sapply(p,prob_win)


# Plot the probability 'p' on the x-axis and 'Pr' on the y-axis.
plot(p,Pr)
```


#### **Exercise 4. A and B play a series - part 2**  
Repeat the previous exercise, but now keep the probability that team A wins fixed at p <- 0.75 and compute the probability for different series lengths. For example, wins in best of 1 game, 3 games, 5 games, and so on through a series that lasts 25 games.

```{r}
# Given a value 'p', the probability of winning the series for the underdog team B can be computed with the following function based on a Monte Carlo simulation:
prob_win <- function(N, p=0.75){
      B <- 10000
      result <- replicate(B, {
        b_win <- sample(c(1,0), N, replace = TRUE, prob = c(1-p, p))
        sum(b_win)>=(N+1)/2
        })
      mean(result)
    }

# Assign the variable 'N' as the vector of series lengths. Use only odd numbers ranging from 1 to 25 games.
N<-seq(1,25,2)

# Apply the 'prob_win' function across the vector of series lengths to determine the probability that team B will win. Call this object `Pr`.
Pr<-sapply(N,prob_win)

# Plot the number of games in the series 'N' on the x-axis and 'Pr' on the y-axis.
plot(N,Pr)
```

