---
title: "3 - Random Variables, Sampling Models, and the Central Limit Theorem"
author: "Sajini Arumugam"
output: html_document
weight: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Random Variables and Sampling Models

Exercise 1. American Roulette probabilities
An American roulette wheel has 18 red, 18 black, and 2 green pockets. Each red and black pocket is associated with a number from 1 to 36. The two remaining green slots feature "0" and "00". Players place bets on which pocket they think a ball will land in after the wheel is spun. Players can bet on a specific number (0, 00, 1-36) or color (red, black, or green).

What are the chances that the ball lands in a green pocket?

```{r}
# The variables `green`, `black`, and `red` contain the number of pockets for each color
green <- 2
black <- 18
red <- 18

# Assign a variable `p_green` as the probability of the ball landing in a green pocket
p_green<-green/(green+black+red)

# Print the variable `p_green` to the console
p_green
```
<br/>

#### **Exercise 2. American Roulette payout**  
In American roulette, the payout for winning on green is $17. This means that if you bet $1 and it lands on green, you get $17 as a prize.

Create a model to predict your winnings from betting on green one time.

```{r}
# Use the `set.seed` function to make sure your answer matches the expected result after random sampling.
set.seed(1)

# The variables 'green', 'black', and 'red' contain the number of pockets for each color
green <- 2
black <- 18
red <- 18

# Assign a variable `p_green` as the probability of the ball landing in a green pocket
p_green <- green / (green+black+red)

# Assign a variable `p_not_green` as the probability of the ball not landing in a green pocket
p_not_green<-(1-p_green)

# Create a model to predict the random variable `X`, your winnings from betting on green. Sample one time.

X <-sample(c(17,-1),1,replace = TRUE,prob=c(p_green,p_not_green))

# Print the value of `X` to the console

X
```
<br/>

#### **Exercise 3. American Roulette expected value**  
In American roulette, the payout for winning on green is $17. This means that if you bet $1 and it lands on green, you get $17 as a prize.In the previous exercise, you created a model to predict your winnings from betting on green.

Now, compute the expected value of X, the random variable you generated previously.

```{r}
# The variables 'green', 'black', and 'red' contain the number of pockets for each color
green <- 2
black <- 18
red <- 18

# Assign a variable `p_green` as the probability of the ball landing in a green pocket
p_green <- green / (green+black+red)

# Assign a variable `p_not_green` as the probability of the ball not landing in a green pocket
p_not_green <- 1-p_green

# Calculate the expected outcome if you win $17 if the ball lands on green and you lose $1 if the ball doesn't land on green

p_green *17 + p_not_green * (-1)
```
<br/>

#### **Exercise 4. American Roulette standard error**  
The standard error of a random variable X tells us the difference between a random variable and its expected value. You calculated a random variable X in exercise 2 and the expected value of that random variable in exercise 3.

Now, compute the standard error of that random variable, which represents a single outcome after one spin of the roulette wheel.

```{r}
# The variables 'green', 'black', and 'red' contain the number of pockets for each color
green <- 2
black <- 18
red <- 18

# Assign a variable `p_green` as the probability of the ball landing in a green pocket
p_green <- green / (green+black+red)

# Assign a variable `p_not_green` as the probability of the ball not landing in a green pocket
p_not_green <- 1-p_green

# Compute the standard error of the random variable
abs((17 - -1))*sqrt(p_green*p_not_green)
```
<br/>

#### **Exercise 5. American Roulette sum of winnings**  
You modeled the outcome of a single spin of the roulette wheel, X, in exercise 2.

Now create a random variable S that sums your winnings after betting on green 1,000 times.

```{r}
# The variables 'green', 'black', and 'red' contain the number of pockets for each color
green <- 2
black <- 18
red <- 18

# Assign a variable `p_green` as the probability of the ball landing in a green pocket
p_green <- green / (green+black+red)

# Assign a variable `p_not_green` as the probability of the ball not landing in a green pocket
p_not_green <- 1-p_green

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# Define the number of bets using the variable 'n'

n<-1000
# Create a vector called 'X' that contains the outcomes of 1000 samples
X<-sample(c(17,-1), size=n, replace=TRUE, prob=c(p_green,p_not_green))

# Assign the sum of all 1000 outcomes to the variable 'S'
S<-sum(X)

# Print the value of 'S' to the console
S
```
<br/>


#### **Exercise 6. American Roulette winnings expected value**  
In the previous exercise, you generated a vector of random outcomes, S, after betting on green 1,000 times.

What is the expected value of S?

```{r}
# The variables 'green', 'black', and 'red' contain the number of pockets for each color
green <- 2
black <- 18
red <- 18

# Assign a variable `p_green` as the probability of the ball landing in a green pocket
p_green <- green / (green+black+red)

# Assign a variable `p_not_green` as the probability of the ball not landing in a green pocket
p_not_green <- 1-p_green

# Define the number of bets using the variable 'n'
n <- 1000

# Calculate the expected outcome of 1,000 spins if you win $17 when the ball lands on green and you lose $1 when the ball doesn't land on green
n*(p_green *17 + p_not_green * (-1))
```
<br/>


#### **Exercise 7. American Roulette winnings expected value**  
You generated the expected value of S, the outcomes of 1,000 bets that the ball lands in the green pocket, in the previous exercise.

What is the standard error of S?

```{r}
# The variables 'green', 'black', and 'red' contain the number of pockets for each color
green <- 2
black <- 18
red <- 18

# Assign a variable `p_green` as the probability of the ball landing in a green pocket
p_green <- green / (green+black+red)

# Assign a variable `p_not_green` as the probability of the ball not landing in a green pocket
p_not_green <- 1-p_green

# Define the number of bets using the variable 'n'
n <- 1000

# Compute the standard error of the sum of 1,000 outcomes
sqrt(n)*(abs((17 - -1))*sqrt(p_green*p_not_green))
```
<br/>

***


### Assessment: The Central Limit Theorem 


#### **Exercise 1. American Roulette probability of winning money**  
The exercises in the previous chapter explored winnings in American roulette. In this chapter of exercises, we will continue with the roulette example and add in the Central Limit Theorem.

In the previous chapter of exercises, you created a random variable S that is the sum of your winnings after betting on green a number of times in American Roulette.

What is the probability that you end up winning money if you bet on green 100 times?

```{r}
# Assign a variable `p_green` as the probability of the ball landing in a green pocket
p_green <- 2 / 38

# Assign a variable `p_not_green` as the probability of the ball not landing in a green pocket
p_not_green <- 1-p_green

# Define the number of bets using the variable 'n'
n <- 100

# Calculate 'avg', the expected outcome of 100 spins if you win $17 when the ball lands on green and you lose $1 when the ball doesn't land on green
avg <- n * (17*p_green + -1*p_not_green)

# Compute 'se', the standard error of the sum of 100 outcomes
se <- sqrt(n) * (17 - -1)*sqrt(p_green*p_not_green)
se
avg
# Using the expected value 'avg' and standard error 'se', compute the probability that you win money betting on green 100 times.
1-pnorm(0,avg,se)
```
<br/>


#### **Exercise 2. American Roulette Monte Carlo simulation**  
Create a Monte Carlo simulation that generates 10,000 outcomes of S, the sum of 100 bets.

Compute the average and standard deviation of the resulting list and compare them to the expected value (-5.263158) and standard error (40.19344) for S that you calculated previously.


```{r}
# Assign a variable `p_green` as the probability of the ball landing in a green pocket
p_green <- 2 / 38

# Assign a variable `p_not_green` as the probability of the ball not landing in a green pocket
p_not_green <- 1-p_green

# Define the number of bets using the variable 'n'
n <- 100

# The variable `B` specifies the number of times we want the simulation to run. Let's run the Monte Carlo simulation 10,000 times.
B <- 10000

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling.
set.seed(1)

# Create an object called `S` that replicates the sample code for `B` iterations and sums the outcomes.
S<-replicate(B,{
  X<-sample(c(17,-1),size=n,replace=TRUE, prob=c(p_green,p_not_green))
  sum(X)
})



# Compute the average value for 'S'
mean(S)

# Calculate the standard deviation of 'S'
sd(S)

```
<br/>

#### **Exercise 3. American Roulette Monte Carlo vs CLT**  
In this chapter, you calculated the probability of winning money in American roulette using the CLT.

Now, calculate the probability of winning money from the Monte Carlo simulation. The Monte Carlo simulation from the previous exercise has already been pre-run for you, resulting in the variable S that contains a list of 10,000 simulated outcomes.

```{r}
# Calculate the proportion of outcomes in the vector `S` that exceed $0
mean(S>0)
```
<br/>


#### **Exercise 4. American Roulette Monte Carlo vs CLT comparison**  
The Monte Carlo result and the CLT approximation for the probability of losing money after 100 bets are close, but not that close. What could account for this?

- 10,000 simulations is not enough. If we do more, the estimates will match.
- The CLT does not work as well when the probability of success is small.   [*]
- The difference is within rounding error.

<br/>


#### **Exercise 5. American Roulette average winnings per bet**  
Now create a random variable Y that contains your average winnings per bet after betting on green 10,000 times.

```{r}
# Use the `set.seed` function to make sure your answer matches the expected result after random sampling.
set.seed(1)

# Define the number of bets using the variable 'n'
n <- 10000

# Assign a variable `p_green` as the probability of the ball landing in a green pocket
p_green <- 2 / 38

# Assign a variable `p_not_green` as the probability of the ball not landing in a green pocket
p_not_green <- 1 - p_green

# Create a vector called `X` that contains the outcomes of `n` bets
X<-sample(c(17,-1), size=n, replace=TRUE, prob=c(p_green,p_not_green))

# Define a variable `Y` that contains the mean outcome per bet. Print this mean to the console.
Y<-mean(X)
Y
```

<br/>

#### **Exercise 6. American Roulette per bet expected value**  
What is the expected value of Y, the average outcome per bet after betting on green 10,000 times?

```{r}
# Assign a variable `p_green` as the probability of the ball landing in a green pocket
p_green <- 2 / 38

# Assign a variable `p_not_green` as the probability of the ball not landing in a green pocket
p_not_green <- 1 - p_green

# Calculate the expected outcome of `Y`, the mean outcome per bet in 10,000 bets
n<-10000
Y<-(p_green *17 + p_not_green * (-1))
Y

```

<br/>

#### **Exercise 7. American Roulette per bet standard error**  
What is the standard error of Y, the average result of 10,000 spins?

```{r}
# Define the number of bets using the variable 'n'
n <- 10000

# Assign a variable `p_green` as the probability of the ball landing in a green pocket
p_green <- 2 / 38

# Assign a variable `p_not_green` as the probability of the ball not landing in a green pocket
p_not_green <- 1 - p_green

# Compute the standard error of 'Y', the mean outcome per bet from 10,000 bets.
Y<-abs((17 - (-1))*sqrt(p_green*p_not_green) / sqrt(n))
Y
```
<br/>

#### **Exercise 8. American Roulette winnings per game are positive**  
What is the probability that your winnings are positive after betting on green 10,000 times?

```{r}
# We defined the average using the following code
avg <- 17*p_green + -1*p_not_green

# We defined standard error using this equation
se <- 1/sqrt(n) * (17 - -1)*sqrt(p_green*p_not_green)

# Given this average and standard error, determine the probability of winning more than $0. Print the result to the console.

1-pnorm(0,avg,se)
```
<br/>

#### **Exercise 9. American Roulette Monte Carlo again**  
Create a Monte Carlo simulation that generates 10,000 outcomes of S, the average outcome from 10,000 bets on green.

Compute the average and standard deviation of the resulting list to confirm the results from previous exercises using the Central Limit Theorem.


```{r}
## Make sure you fully follow instructions, including printing values to the console and correctly running the `replicate` loop. If not, you may encounter "Session Expired" errors.

# The variable `n` specifies the number of independent bets on green
n <- 10000

# The variable `B` specifies the number of times we want the simulation to run
B <- 10000

# Use the `set.seed` function to make sure your answer matches the expected result after random number generation
set.seed(1)

# Generate a vector `S` that contains the the average outcomes of 10,000 bets modeled 10,000 times
S<-replicate(B,{
  X<-sample(c(17,-1),size=n,replace=TRUE, prob=c(p_green,p_not_green))
  mean(X)
})


# Compute the average of `S`
mean(S)

# Compute the standard deviation of `S`
sd(S)
```
<br/>

#### **Exercise 10. American Roulette comparison**  
In a previous exercise, you found the probability of winning more than $0 after betting on green 10,000 times using the Central Limit Theorem. Then, you used a Monte Carlo simulation to model the average result of betting on green 10,000 times over 10,000 simulated series of bets.

What is the probability of winning more than $0 as estimated by your Monte Carlo simulation? The code to generate the vector S that contains the the average outcomes of 10,000 bets modeled 10,000 times has already been run for you.

```{r}
# Compute the proportion of outcomes in the vector 'S' where you won more than $0
mean(S>0)

```
<br/>

#### **Exercise 11. American Roulette comparison analysis**  
The Monte Carlo result and the CLT approximation are now much closer than when we calculated the probability of winning for 100 bets on green. What could account for this difference?

- We are now computing averages instead of sums.
- 10,000 Monte Carlo simulations was not enough to provide a good estimate.
- The CLT works better when the sample size is larger.    [*]
- It is not closer. The difference is within rounding error.
<br/>


### Questions 1 and 2: SAT testing

The SAT is a standardized college admissions test used in the United States. The following two multi-part questions will ask you some questions about SAT testing.

This is a 6-part question asking you to determine some probabilities of what happens when a student guessed for all of their answers on the SAT. Use the information below to inform your answers for the following questions.

An old version of the SAT college entrance exam had a -0.25 point penalty for every incorrect answer and awarded 1 point for a correct answer. The quantitative test consisted of 44 multiple-choice questions each with 5 answer choices. Suppose a student chooses answers by guessing for all questions on the test.


#### **Question 1a**  

What is the probability of guessing correctly for one question?

```{r}
p <- 1/5 # one correct choice of 5 options
p
```
<br/>

#### **Question 1b**  

What is the expected value of points for guessing on one question?
```{r}
a <- 1
b <- -0.25
mu <- a*p + b*(1-p)
mu
```
<br/>

#### **Question 1c**  

What is the expected score of guessing on all 44 questions?

```{r}
n <- 44
n*mu
```
<br/>


#### **Question 1d**  

What is the standard error of guessing on all 44 questions?

```{r}
sigma <- sqrt(n) * abs(b-a) * sqrt(p*(1-p))
sigma
```
<br/>


#### **Question 1e**  

Use the Central Limit Theorem to determine the probability that a guessing student scores 8 points or higher on the test.

```{r}
1-pnorm(8, mu, sigma)
```

#### **Question 1f**  

Set the seed to 21, then run a Monte Carlo simulation of 10,000 students guessing on the test.

(IMPORTANT! If you use R 3.6 or later, you will need to use the command set.seed(x, sample.kind = "Rounding") instead of set.seed(x). Your R version will be printed at the top of the Console window when you start RStudio.)

What is the probability that a guessing student scores 8 points or higher?

```{r}
set.seed(21, sample.kind = "Rounding")
B <- 10000
n <- 44
p <- 0.2
tests <- replicate(B, {
  X <- sample(c(1, -0.25), n, replace = TRUE, prob = c(p, 1-p))
  sum(X)
})
mean(tests >= 8)
```
<br/>


#### **Question 2a**  

Suppose that the number of multiple choice options is 4 and that there is no penalty for guessing - that is, an incorrect question gives a score of 0.

What is the expected value of the score when guessing on this new test?

```{r}
p <- 1/4
a <- 1
b <- 0
n <- 44
mu <- n * a*p + b*(1-p)
mu
```
<br/>


#### **Question 2b**  

Using the normal approximation, what is the estimated probability of scoring over 30 when guessing?

```{r}
sigma <- sqrt(n) * abs(b-a) * sqrt(p*(1-p))
1-pnorm(30, mu, sigma)
```
<br/>

#### **Question 2c**  

Consider a range of correct answer probabilities p <- seq(0.25, 0.95, 0.05) representing a range of student skills.

What is the lowest p such that the probability of scoring over 35 exceeds 80%?

```{r}
p <- seq(0.25, 0.95, 0.05)
exp_val <- sapply(p, function(x){
  mu <- n * a*x + b*(1-x)
  sigma <- sqrt(n) * abs(b-a) * sqrt(x*(1-x))
  1-pnorm(35, mu, sigma)
})

min(p[which(exp_val > 0.8)])
```
<br/>

### Question 3: Betting on Roulette

A casino offers a House Special bet on roulette, which is a bet on five pockets (00, 0, 1, 2, 3) out of 38 total pockets. The bet pays out 6 to 1. In other words, a losing bet yields -$1 and a successful bet yields $6. A gambler wants to know the chance of losing money if he places 500 bets on the roulette House Special.

The following 7-part question asks you to do some calculations related to this scenario.

*** 

#### **Question 3a**  

What is the expected value of the payout for one bet?

```{r}
p <- 5/38
a <- 6
b <- -1
mu <- a*p + b*(1-p)
mu
```
<br/>

#### **Question 3b**  

What is the standard error of the payout for one bet?

```{r}
sigma <- abs(b-a) * sqrt(p*(1-p))
sigma
```
<br/>

#### **Question 3c**  

What is the expected value of the average payout over 500 bets?

```mu
```

#### **Question 3d**  

What is the standard error of the average payout over 500 bets?

```{r}
n <- 500
sigma/sqrt(n)
```
<br/>

#### **Question 3e**  

What is the expected value of the sum of 500 bets?

```
n*mu
```
<br/>

#### **Question 3f**  

What is the standard error of the sum of 500 bets?

```{r}
sqrt(n) * sigma
```
<br/>

#### **Question 3g**  

Use pnorm with the expected value of the sum and standard error of the sum to calculate the probability of losing money over 500 bets, Pr(X≤0).

```{r}
pnorm(0, n*mu, sqrt(n)*sigma)
```

