---
title: "4 - The Big Short"
author: "Sajini Arumugam"
output: html_document
weight: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### DataCamp Assessment: The Big Short

#### **Exercise 1. Bank earnings**  
Say you manage a bank that gives out 10,000 loans. The default rate is 0.03 and you lose $200,000 in each foreclosure.

Create a random variable S that contains the earnings of your bank. Calculate the total amount of money lost in this scenario.

```{r}
# Assign the number of loans to the variable `n`
n <- 10000

# Assign the loss per foreclosure to the variable `loss_per_foreclosure`
loss_per_foreclosure <- -200000

# Assign the probability of default to the variable `p_default`
p_default <- 0.03

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# Generate a vector called `defaults` that contains the default outcomes of `n` loans
defaults <- sample( c(0,1), n, replace = TRUE, prob=c(1-p_default,p_default))


# Generate `S`, the total amount of money lost across all foreclosures. Print the value to the console.
S<-sum(defaults * loss_per_foreclosure)
S

```
<br/>

#### **Exercise 2. Bank earnings Monte Carlo**  
Run a Monte Carlo simulation with 10,000 outcomes for S, the sum of losses over 10,000 loans. Make a histogram of the results.

```{r}
# Assign the number of loans to the variable `n`
n <- 10000

# Assign the loss per foreclosure to the variable `loss_per_foreclosure`
loss_per_foreclosure <- -200000

# Assign the probability of default to the variable `p_default`
p_default <- 0.03

# Use the `set.seed` function to make sure your answer matches the expected result after random sampling
set.seed(1)

# The variable `B` specifies the number of times we want the simulation to run
B <- 10000

# Generate a list of summed losses 'S'. Replicate the code from the previous exercise over 'B' iterations to generate a list of summed losses for 'n' loans.  Ignore any warnings for now.
S<-replicate (B,{
  defaults <- sample( c(0,1), n, replace = TRUE, prob=c(1-p_default,p_default))
  sum(defaults*loss_per_foreclosure)
  })

hist(S)

```
<br/>

#### **Exercise 3. Bank earnings expected value**  
What is the expected value of S, the sum of losses over 10,000 loans? For now, assume a bank makes no money if the loan is paid.

```{r}
# Assign the number of loans to the variable `n`
n <- 10000

# Assign the loss per foreclosure to the variable `loss_per_foreclosure`
loss_per_foreclosure <- -200000

# Assign the probability of default to the variable `p_default`
p_default <- 0.03

# Calculate the expected loss due to default out of 10,000 loans
n*((p_default*loss_per_foreclosure)+(1-p_default)*0)
```
<br/>

#### **Exercise 4. Bank earnings standard error**  
What is the standard error of S?

```{r}
# Assign the number of loans to the variable `n`
n <- 10000

# Assign the loss per foreclosure to the variable `loss_per_foreclosure`
loss_per_foreclosure <- -200000

# Assign the probability of default to the variable `p_default`
p_default <- 0.03

# Compute the standard error of the sum of 10,000 loans
sqrt(n)*abs(loss_per_foreclosure)*sqrt(p_default*(1-p_default))
```
<br/>

#### **Exercise 5. Bank earnings interest rate - 1**  
So far, we've been assuming that we make no money when people pay their loans and we lose a lot of money when people default on their loans. Assume we give out loans for $180,000. How much money do we need to make when people pay their loans so that our net loss is $0?

In other words, what interest rate do we need to charge in order to not lose money?

```{r}
# Assign the loss per foreclosure to the variable `loss_per_foreclosure`
loss_per_foreclosure <- -200000

# Assign the probability of default to the variable `p_default`
p_default <- 0.03

# Assign a variable `x` as the total amount necessary to have an expected outcome of $0
x<--(loss_per_foreclosure*p_default)/(1-p_default)

# Convert `x` to a rate, given that the loan amount is $180,000. Print this value to the console.

x/180000
```
<br/>

#### **Exercise 6. Bank earnings interest rate - 2**  
With the interest rate calculated in the last example, we still lose money 50% of the time. What should the interest rate be so that the chance of losing money is 1 in 20?

In math notation, what should the interest rate be so that Pr(S<0)=0.05?

Remember that we can add a constant to both sides of the equation to get:
Pr(S−E[S]SE[S]<−E[S]SE[S])
which is

Pr(Z<−[lp+x(1−p)]n(x−l)np(1−p)−−−−−−−−√)=0.05
Let z = qnorm(0.05) give us the value of z for which:
Pr(Z≤z)=0.05

```{r}
# Assign the number of loans to the variable `n`
n <- 10000

# Assign the loss per foreclosure to the variable `loss_per_foreclosure`
loss_per_foreclosure <- -200000

# Assign the probability of default to the variable `p_default`
p_default <- 0.03
```
<br/>

#### **Exercise 7. Bank earnings - minimize money loss**  
The bank wants to minimize the probability of losing money. Which of the following achieves their goal without making interest rates go up?

- A smaller pool of loans
- A larger probability of default
- A reduced default rate    [*]
- A larger cost per loan default





